---
title: "journalist_diversity"
author: "Natalie Davidson"
date: "10/19/2021"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

require(data.table)
require(here)
require(ggplot2)
require(caret)
require(stringr)
require(ggrepel)
options(ggrepel.max.overlaps = Inf)


proj_dir = here()
source(file.path(proj_dir, "/analysis_scripts/analysis_utils.R"))
source(file.path(proj_dir, "/utils/plotting_utils.R"))
source(file.path(proj_dir, "/utils/scraper_processing_utils.R"))
source(file.path(proj_dir, "/process_doi_data/springer_scripts/springer_scrape_utils.R"))

RERUN_BOOTSTRAP=FALSE

```

## Nature News comparison of journalists

This document is an analysis of the extracted journalists names and their gender and name origin predictions.
It calculates boostrap estimations over all Nature news articles and by article type.



The **setting + helper functions** to generate the plots are here:

1. plotting related functions: `/utils/plotting_utils.R`

2. reading + data processing related functions: `/utils/scraper_processing_utils.R` and `/analysis_scripts/analysis_utils.R`

3. nature research article and springer specific data processing functions: `/process_doi_data/springer_scripts/springer_scrape_utils.R`


## Read in data


### reading in the quote data

```{r read_json_data, fig.align='center', echo=TRUE, warning=FALSE, message=F}


# read in the scraped citations from nature news articles for each year
journo_dir = file.path(proj_dir, "/data/journalist_data/")
all_authors = NA


tab_res_files = list.files(journo_dir, pattern=".tsv", full.names = TRUE)

all_authors = NA
for(curr_file in tab_res_files){

    print(curr_file)
    
    curr_type = substr(basename(curr_file), 19, nchar(basename(curr_file))-4)

    # skip empty files
    if(file.info(curr_file)$size == 0){
        next
    }
    
    res_df = data.frame(fread(curr_file))
    res_df = subset(res_df, authors != "")
    res_df$type = curr_type

    # make df
    authors_df = res_df[,c("file_id", "year", "authors", "type")]
    all_authors = rbind(all_authors, authors_df)

}

all_authors = all_authors[-1,]

# format file_id into a doi
all_authors$doi = paste("doi:10.1038/", all_authors$file_id, sep="")


# split up all the journalist names
author_df_split = separate_rows(all_authors, "authors",sep = "; ")
author_df_split = subset(author_df_split, authors != "")


# remove the article types we are not interested in
all_authors = subset(all_authors, !type %in% c("career-column", "news-and-views"))

# plot number of articles scraped
ggplot(unique(all_authors[,c("file_id", "year")]), aes(x=as.factor(year))) +
    geom_bar() + theme_bw() +
    xlab("Year of Article") + ylab("# articles") +
        ggtitle("# Nature News Articles with Author Over Time")



```


### reading in the first and last author data

First, read in gender data

```{r read_gender, fig.align='center', echo=TRUE, warning=FALSE, message=F}

# read in the scraped nature data
journo_gender = file.path(proj_dir,
                    "/data/author_data/nature_journo_gender.tsv")
gender_res = data.frame(fread(journo_gender))
gender_res$est_gender = gender_res$gender
gender_res = merge(gender_res, all_authors[,c("type", "file_id", "year", "doi")])

# remove the article types we are not interested in
gender_res = subset(gender_res, !type %in% c("career-column", "news-and-views"))

# remove quotes where no gender could be estimated
gender_res = gender_res[gender_res$est_gender %in% c("FEMALE", "MALE"), ]
gender_res = subset(gender_res, !is.na(est_gender))


```

Now, read in name-origin data


```{r read_name_origin, fig.align='center', echo=TRUE, warning=FALSE, message=F}

# now read in the BG data
name_pred_file = file.path(proj_dir, 
                         "/data/author_data/all_author_fullname_pred.tsv")
name_info_file = file.path(proj_dir, 
                         "/data/author_data/all_author_fullname.tsv")
journo_name_df = read_name_origin(name_pred_file, name_info_file)
journo_name_df$name_origin[journo_name_df$name_origin == "Jewish"] = "Hebrew"
journo_name_df = merge(journo_name_df, all_authors[,c("type", "file_id", "year", "doi")])

# remove the article types we are not interested in
journo_name_df = subset(journo_name_df, !type %in% c("career-column", "news-and-views"))

```

## Calculate bootstrap estimates

### gender

```{r bootstrap_est_gender, echo=TRUE, warning=FALSE, message=F}

if(RERUN_BOOTSTRAP){
        
    
    #### journo data
    journo_male_prop_df = compute_bootstrap_gender(gender_res, 
                                               year_col_id = "year", 
                                               article_col_id = "file_id",
                                               conf_int=0.95)
    journo_male_prop_df$corpus = "journo_m"

    journo_female_prop_df = compute_bootstrap_gender(gender_res, 
                                               year_col_id = "year", 
                                               article_col_id = "file_id",
                                               conf_int=0.95,
                                               gender_search_str="FEMALE")
    journo_female_prop_df$corpus = "journo_f"
        
    
    #### journo data broken down by article type
    get_subboot_gender <- function(type_id, type_names, in_df, gender_search_str){
        bootstrap_res = compute_bootstrap_gender(subset(in_df, type %in% type_id), 
                                               year_col_id = "year", 
                                               article_col_id = "file_id",
                                               conf_int=0.95,
                                               gender_search_str)
        bootstrap_res$corpus = type_names
        return(bootstrap_res)
    
    }
    
    type_df_m = NA
    for(curr_type in unique(gender_res$type)){
        res = get_subboot_gender(curr_type, curr_type, gender_res, gender_search_str="MALE")
        type_df_m = rbind(type_df_m, res)
    }
    type_df_m = type_df_m[-1,]
    
    type_df_f = NA
    for(curr_type in unique(gender_res$type)){
        res = get_subboot_gender(curr_type, curr_type, gender_res, gender_search_str="FEMALE")
        type_df_f = rbind(type_df_f, res)
    }
    type_df_f = type_df_f[-1,]
        
    all_bootstrap_file = file.path(proj_dir,
                            "/figure_notebooks/supp_analyze_journalist_data/tmp_files/journo_gender.RData")
    save(journo_male_prop_df, journo_female_prop_df, type_df_m, type_df_f, file = all_bootstrap_file)
    
    
}else{
    all_bootstrap_file = file.path(proj_dir,
                            "/figure_notebooks/supp_analyze_journalist_data/tmp_files/journo_gender.RData")
    load(all_bootstrap_file)
}
journo_male_prop_df$corpus = "male_journalist"
journo_female_prop_df$corpus = "female_journalist"

journo_gender_prop_df = rbind(journo_male_prop_df, journo_female_prop_df)


```

### name origin


```{r bootstrap_est_name_origin, echo=TRUE, warning=FALSE, message=F}

get_subboot_name_origin <- function(origin_id, curr_corpus, in_df, bootstrap_col_id="doi"){
    bootstrap_res = compute_bootstrap_location(subset(in_df, 
                                                      corpus==curr_corpus), 
                                              year_col_id = "year", 
                                              article_col_id = bootstrap_col_id, 
                                              country_col_id = "name_origin",
                                              country_agg = origin_id, 
                                              conf_int = 0.95)
    bootstrap_res$name_origin = origin_id
    
    # add a label for plotting later
    bootstrap_res$label[bootstrap_res$year == 2020] = 
        bootstrap_res$name_origin[bootstrap_res$year == 2020]
        
    return(bootstrap_res)

}
if(RERUN_BOOTSTRAP){
        
    journo_origin_df = NA
    for(curr_origin in unique(journo_name_df$name_origin)){
        print(curr_origin)
        res = get_subboot_name_origin(curr_origin, 
                          curr_corpus="nature_journo",
                          journo_name_df)
        journo_origin_df = rbind(journo_origin_df, res)
    }
    journo_origin_df = journo_origin_df[-1,]
    
    all_bootstrap_file = file.path(proj_dir,
                            "/figure_notebooks/supp_analyze_journalist_data/tmp_files/journo_name_origin.RData")
    save(journo_origin_df, file = all_bootstrap_file)
    
    
}else{
    all_bootstrap_file = file.path(proj_dir,
                            "/figure_notebooks/supp_analyze_journalist_data/tmp_files/journo_name_origin.RData")
    load(all_bootstrap_file)
}



```

## Make Plots

```{r make_citation_gg, echo=TRUE, warning=FALSE, message=F}

#### write out tables to show the raw frequencies
journo_f_df = subset(gender_res, est_gender=="FEMALE")
journo_f_tab = table(journo_f_df$type, journo_f_df$year)

journo_m_df = subset(gender_res, est_gender=="MALE")
journo_m_tab = table(journo_m_df$type, journo_m_df$year)


#### plot the overview of gender
journo_gender_gg = 
    ggplot(journo_gender_prop_df, 
      aes(x=as.numeric(year), y=mean,
                          ymin=bottom_CI, ymax=top_CI,
                          fill=corpus)) +
    geom_point() + geom_ribbon(alpha=0.5) + geom_line(alpha=0.5) + theme_bw() + 
    xlab("Year of Article") + ylab("Male Percentage") +
    ggtitle("Est. Proportion Written Articles by Gender in Nature") + 
    ylim(c(0, 1)) +
    geom_hline(yintercept=0.5, color="red") +
    scale_fill_manual(values=c("#E69F00", "#56B4E9")) +
    theme(legend.position="bottom")

ggsave(file.path(proj_dir,  "/figure_notebooks/supp_analyze_journalist_data/tmp_files/journo_gender_gg.pdf"),
       journo_gender_gg, width = 6, height = 5, units = "in", device = "pdf")

journo_gender_type_m_gg = 
    ggplot(type_df_m, 
      aes(x=as.numeric(year), y=mean,
                          ymin=bottom_CI, ymax=top_CI,
                          fill=corpus)) +
    geom_point() + geom_ribbon(alpha=0.5) + geom_line(alpha=0.5) + theme_bw() + 
    xlab("Year of Article") + ylab("Male Percentage") +
    ggtitle("Est. Proportion Male Written Articles in Nature") + 
    ylim(c(0, 1)) +
    geom_hline(yintercept=0.5, color="red") +
    theme(legend.position="bottom")

ggsave(file.path(proj_dir,  "/figure_notebooks/supp_analyze_journalist_data/tmp_files/journo_gender_type_m_gg.pdf"),
       journo_gender_type_m_gg, width = 6, height = 5, units = "in", device = "pdf")


journo_gender_type_f_gg = 
    ggplot(type_df_f, 
      aes(x=as.numeric(year), y=mean,
                          ymin=bottom_CI, ymax=top_CI,
                          fill=corpus)) +
    geom_point() + geom_ribbon(alpha=0.5) + geom_line(alpha=0.5) + theme_bw() + 
    xlab("Year of Article") + ylab("Male Percentage") +
    ggtitle("Est. Proportion Female Written Articles in Nature") + 
    ylim(c(0, 1)) +
    geom_hline(yintercept=0.5, color="red") +
    theme(legend.position="bottom")

ggsave(file.path(proj_dir,  "/figure_notebooks/supp_analyze_journalist_data/tmp_files/journo_gender_type_f_gg.pdf"),
       journo_gender_type_f_gg, width = 6, height = 5, units = "in", device = "pdf")



#### plot the overview of name origin
journo_name_origin_gg = ggplot(journo_origin_df, aes(x=as.numeric(year), y=mean,
                          ymin=bottom_CI, ymax=top_CI,
                          fill=name_origin, label=label)) +
    geom_point() + geom_ribbon(alpha=0.5) + geom_line(alpha=0.5) +
    theme_bw() + geom_text_repel() + xlim(c(2005,2021))  +
    xlab("Year of Article") + ylab("Proportion of Journalist Citations") +
    ggtitle("Est. proportion of articles written by specific name origin ") + 
    scale_fill_brewer(palette="Set2") +
    theme(legend.position = "none")

ggsave(file.path(proj_dir,  "/figure_notebooks/supp_analyze_journalist_data/tmp_files/journo_name_origin_gg.pdf"),
       journo_name_origin_gg, width = 7, height = 5, units = "in", device = "pdf")

journo_gender_gg


knitr::kable(journo_f_tab, format = "pipe", 
             caption = "Number of Articles written by predicted female journalist")

knitr::kable(journo_m_tab, format = "pipe", 
             caption = "Number of Articles written by predicted male journalist")


journo_gender_type_m_gg
journo_gender_type_f_gg

journo_name_tab = table(journo_name_df$name_origin, journo_name_df$year)
knitr::kable(journo_name_tab, format = "pipe", 
             caption = "Number of Articles written by a journalist with predicted name-origin ")

journo_name_origin_gg
```
