---
title: "overview_fig"
author: "Natalie Davidson"
date: "4/30/2021"
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

require(data.table)
require(here)
require(ggplot2)
require(magick)
require(pdftools)

proj_dir = here()
source(file.path(proj_dir, "/utils/scraper_processing_utils.R"))
source(file.path(proj_dir, "/utils/plotting_utils.R"))
source(file.path(proj_dir, "/process_doi_data/springer_scripts/springer_scrape_utils.R"))

```


## Overview

This notebook generates figure 1 and supplemental figure 1.


The **data** it uses to build the plots are here:

1. nature news scraped data: `/data/scraped_data/downloads/`

2. nature research articles and letters author data: `/data/author_data/downloads/`


The **pdfs** included in the plots are here:

1. `/figure_notebooks/illustrator_pdfs/`

The **setting + helper functions** to generate the plots are here:

1. plotting related functions: `/utils/plotting_utils.R`

2. reading + data processing related functions: `/utils/scraper_processing_utils.R`

3. nautre research article and springer specific data processing functions: `/process_doi_data/springer_scripts/springer_scrape_utils.R`




## Read in the data

### Read in the scraped DOIs

```{r read_nature_news, echo=TRUE, warning=FALSE, message=F}


# read in the scraped news articles for each year
news_scraped_dir = file.path(proj_dir,
                    "/data/scraped_data/downloads/")
news_scraped_dir_files = list.files(news_scraped_dir, full.names = T)

news_df = NA
for(curr_file in news_scraped_dir_files){
    
    curr_df = read_json(curr_file)
    
    # if the json file was empty, skip
    if(all(is.na(curr_df))){
        next
    }
    
    # get the year form the file name
    file_name_year = substring(basename(curr_file), 
                            15, 18)
    
    # get the news article type from the file name
    file_name_type = substring(basename(curr_file), 
                            20, nchar(basename(curr_file))-5)
    
    # format the output
    article_ids = curr_df$file_id
    num_articles = length(article_ids)
    curr_info_df = data.frame(year=file_name_year, 
                                type=file_name_type, 
                                file_id=article_ids)
    news_df = rbind(news_df, curr_info_df)
    
}
news_df = news_df[-1,]

head(news_df)

```

### Read in the nature + springer research author information

```{r read_nature_articles, echo=TRUE, warning=FALSE, message=F}

# now read all the nature the research + letters articles
nature_author_dir = file.path(proj_dir,
                    "/data/author_data/downloads/")
nature_author_df = read_nature_author_json_files(nature_author_dir)

head(nature_author_df)

# now read the randomly sampled Springer research articles
springer_author_file = file.path(proj_dir, 
                                "/data/reference_data/springer_bg_author_cache.tsv")
springer_author_df = data.frame(fread(springer_author_file))

head(springer_author_df)


```


### append the three data sources together

```{r append_data_together, echo=TRUE, warning=FALSE, message=F}

nature_author_df$type = "research"
springer_author_df$type = "springer"
springer_author_df$file_id = springer_author_df$doi

col_ids = c("year", "type", "file_id")
full_df = Reduce(rbind, list(nature_author_df[,col_ids], 
                             springer_author_df[,col_ids],
                             news_df[,col_ids]))

```



## Make the Figures

### generate the plots

```{r make_gg, echo=TRUE, warning=FALSE, message=F}

# plot number of articles scraped
full_df$type = factor(full_df$type, levels = ARTICLE_TYPE_FACTOR_ORDER)
news_nature_gg = ggplot(subset(full_df, type != "springer"), 
                        aes(x=as.factor(year), fill=type)) +
                geom_bar() + theme_bw() +
                scale_fill_manual(values=ARTICLE_TYPE_COLOR) +
                xlab("Year of Article") + ylab("Num. Articles") +
                    ggtitle("# Nature News and Research Articles Over Time")

all_gg = ggplot(full_df, aes(x=as.factor(year), fill=type)) +
                geom_bar() + theme_bw() +
                scale_fill_manual(values=ARTICLE_TYPE_COLOR) +
                xlab("Year of Article") + ylab("# articles") +
                    ggtitle("# Springer + Nature News and Research Articles Over Time")
ggsave(file.path(proj_dir, "/figure_notebooks/tmp_files/fig1_tmp/news_nature_gg.png"),
       news_nature_gg, width = 10, height = 4, units = "in")
ggsave(file.path(proj_dir, "/figure_notebooks/tmp_files/fig1_tmp/all_gg.png"),
       all_gg)

```


### format figure 1

```{r make_fig1, echo=TRUE, warning=FALSE, message=F}

plot_overview = image_read_pdf(file.path(proj_dir,
                                  "/figure_notebooks/illustrator_pdfs/nature_news_ex_fig1a.pdf"))
plot_data_diagram = image_read_pdf(file.path(proj_dir,
                                  "/figure_notebooks/illustrator_pdfs/nature_news_ex_fig1b.pdf"))
news_nature_gg = image_read(file.path(proj_dir,
                                  "/figure_notebooks/tmp_files/fig1_tmp/news_nature_gg.png"))


bottom_panel <- image_append(image_scale(c(plot_overview, plot_data_diagram), 500), stack = FALSE)
full_image <- image_append(image_scale(c(bottom_panel, news_nature_gg),3000), 
                           stack = TRUE)

print(full_image)

outfile = file.path(proj_dir,"/figure_notebooks/tmp_files/fig1_tmp/fig1_main.pdf")
image_write(full_image, format = "pdf", outfile)

```

