---
title: "Supp_tables"
author: "Natalie Davidson"
date: "6/15/2021"
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

require(data.table)
require(here)
require(ggplot2)
require(magick)
require(pdftools)
require(stringr)

proj_dir = here()
source(file.path(proj_dir, "/utils/scraper_processing_utils.R"))
source(file.path(proj_dir, "/utils/plotting_utils.R"))
source(file.path(proj_dir, "/analysis_scripts/analysis_utils.R"))
source(file.path(proj_dir, "/process_doi_data/springer_scripts/springer_scrape_utils.R"))

```


## Overview

This notebook generates supplemental tables 1-4


The **data** it uses to build the tables are here:

1. nature news scraped data after coreNLP: `/data/scraped_data/downloads/coreNLP_output*`

2. nature news quotes: `/data/scraped_data/downloads/quote_table_raw*`

3. quoted speakers and name origin prediction: 
    `/data/author_data/all_speaker_fullname_pred.tsv`
    `/data/author_data/all_speaker_fullname.tsv`

4. springer authors: `/data/reference_data/springer_bg_author_cache.tsv`

5. Nature News cited authors in a springer journal: `/data/reference_data/springer_cited_author_cache.tsv`

6. All cited DOIs: `data/doi_data/downloads/*`

7. All authors' names and name origin predictions: `/data/author_data/all_author_fullname_pred.tsv`

8. Nature and Springer Authors' gender predictions: 
    `/data/author_data/nature_author_gender.tsv`
    `/data/author_data/springer_author_gender.tsv`



The **setting + helper functions** to generate the plots are here:

1. plotting related functions: `/utils/plotting_utils.R`

2. reading + data processing related functions: `/utils/scraper_processing_utils.R`

3. nautre research article and springer specific data processing functions: `/process_doi_data/springer_scripts/springer_scrape_utils.R`




## Read in the data


### Read in reference data
```{r read_ref_articles, echo=TRUE, warning=FALSE, message=F}

# read in the scraped news articles for each year
# we will need this later for filtering out articles in columns
# we would like to ignore
news_scraped_dir = file.path(proj_dir,
                    "/data/scraped_data/")
news_scraped_dir_files = list.dirs(news_scraped_dir, full.names = T)
news_scraped_dir_files = grep("coreNLP_output", news_scraped_dir_files, value=T)

news_df = NA
for(curr_dir in news_scraped_dir_files){
    
    curr_files = list.files(curr_dir, full.names = T)

    
    # if the json file was empty, skip
    if(length(curr_files) == 0 ){
        next
    }
    
    # get the year form the file name
    file_name_year = substring(basename(curr_dir), 
                            16, 19)
    
    # get the news article type from the file name
    file_name_type = substring(basename(curr_dir), 
                            21, nchar(basename(curr_dir)))
    
    # format the output
    article_ids = gsub(".txt.json", "", basename(curr_files))
    num_articles = length(article_ids)
    curr_info_df = data.frame(year=file_name_year, 
                                type=file_name_type, 
                                file_id=article_ids)
    news_df = rbind(news_df, curr_info_df)
    
}
news_df = news_df[-1,]

# filter out career column and news-and-views
news_df = subset(news_df, !type %in% c("career-column", "news-and-views"))
head(news_df)

```

### Read in the quotes

```{r read_quotes, echo=TRUE, warning=FALSE, message=F}


# read in the all quotes for all news articles and all years 
full_quote_df = NA
quote_files = list.files(file.path(proj_dir,"/data/scraped_data/", sep=""), full.names = T)
quote_files = grep("quote_table_raw_", quote_files, value=T)
for(quote_file in quote_files){
    
    quote_df = read_corenlp_quote_files(quote_file)
    quote_df$year = str_extract(quote_file, "[1-9][0-9]+") # curr_year
    quote_df$type = substring(basename(quote_file), 
                            22, nchar(basename(quote_file))-4)
    
    full_quote_df = rbind(full_quote_df, quote_df)
}
full_quote_df = full_quote_df[-1,]


# read in the name prediction information for quotes
name_pred_file = file.path(proj_dir, 
                             "/data/author_data/all_speaker_fullname_pred.tsv")
name_info_file = file.path(proj_dir, 
                             "/data/author_data/all_speaker_fullname.tsv")
quote_name_origin_df = read_name_origin(name_pred_file, name_info_file)
quote_name_origin_df = subset(quote_name_origin_df, !type %in% c("career-column", "news-and-views"))


```

### Read in the authorship info (springer, nature, and cited articles)
```{r quotes_gender_pred_table1, echo=TRUE, warning=FALSE, message=F}

# we have 3 source files for author info
# springer background authorship
springer_author_file = file.path(proj_dir, 
                                "/data/reference_data/springer_bg_author_cache.tsv")
springer_author_df = data.frame(fread(springer_author_file))
springer_author_df$file_id = NA

# springer cited authorship
cited_author_file = file.path(proj_dir, 
                                "/data/reference_data/springer_cited_author_cache.tsv")
cited_dois_dir = file.path(proj_dir, "data/doi_data/downloads")

cited_author_df = data.frame(fread(cited_author_file))
cited_author_df = subset(cited_author_df, !is.na(authors))
cited_author_df$pub_year = cited_author_df$year
cited_author_df = subset(cited_author_df, select=-c(year))
cited_doi_df = get_ref_dois(cited_dois_dir)
cited_author_df = merge(cited_author_df, cited_doi_df[,c("doi", "year", "file_id")], by=c("doi"))

# then all the nature articles
nature_dir = file.path(proj_dir, "data/author_data/downloads")
nature_author_df = read_nature_author_json_files(nature_dir)


# now read in the processed data for name prediction
name_pred_file = file.path(proj_dir, 
                         "/data/author_data/all_author_fullname_pred.tsv")
name_info_file = file.path(proj_dir, 
                         "/data/author_data/all_author_fullname.tsv")
origin_cite_name_df = read_name_origin(name_pred_file, name_info_file)
origin_cite_name_df$name_origin[origin_cite_name_df$name_origin == "Jewish"] = "Hebrew"


# seperate out citations from columns by journalists vs scientists
journalist_idx = which(origin_cite_name_df$corpus == "naturenews_citations" &
                         origin_cite_name_df$file_id %in% news_df$file_id)
scientist_idx = which(origin_cite_name_df$corpus == "naturenews_citations" &
                         !origin_cite_name_df$file_id %in% news_df$file_id)
origin_cite_name_df$corpus[journalist_idx] = "citation_journalist"
origin_cite_name_df$corpus[scientist_idx] = "citation_scientist"

# now read in the processed data for gender prediction
springer_gender_file = file.path(proj_dir, 
                         "/data/author_data/springer_author_gender.tsv")
springer_pred_gender_df = data.frame(fread(springer_gender_file))

nature_file = file.path(proj_dir, "/data/author_data/nature_author_gender.tsv")
nature_pred_gender_df = read_gender_files(nature_file)
head(nature_pred_gender_df)


```


## Get Table info

### Gender prediction for quotes
```{r quotes_gender_pred_table1, echo=TRUE, warning=FALSE, message=F}


# filter out career column and news-and-views
full_quote_df_gender = subset(full_quote_df, !type %in% c("career-column", "news-and-views"))
full_quote_df_gender = unique(full_quote_df_gender)

print("Total Quotes")
total_quotes = nrow(full_quote_df_gender)
print(total_quotes)

# remove names with single name and do not have a pronoun
space_idx = grep(" ", full_quote_df_gender$est_speaker)
gendered_pronouns = c("he", "him", "his", "himself",
                    "she", "her", "hers", "herself")
pronoun_idx_canonical = which(full_quote_df_gender$canonical_speaker %in% gendered_pronouns)
pronoun_idx_partial = which(full_quote_df_gender$partial_name %in% gendered_pronouns)
allowed_idx = unique(c(space_idx, pronoun_idx_canonical, pronoun_idx_partial))
full_quote_df_gender = full_quote_df_gender[allowed_idx,]
full_quote_df_gender = unique(full_quote_df_gender)

print("Total quotes with full name or pronoun")
quotes_with_name_or_pronoun = nrow(full_quote_df_gender)
print(quotes_with_name_or_pronoun)


# remove quotes where no gender could be estimated
full_quote_df_gender = subset(full_quote_df_gender, !is.na(est_gender))
full_quote_df_gender = full_quote_df_gender[full_quote_df_gender$est_gender %in% c("FEMALE", "MALE"), ]

print("Total quotes with Gender Prediction")
full_quote_df_gender = unique(full_quote_df_gender)
quotes_with_gender = nrow(full_quote_df_gender)
print(quotes_with_gender)


```



### Name origin prediction for quotes
```{r quotes_origin_pred_table1, echo=TRUE, warning=FALSE, message=F}

# filter out career column and news-and-views
full_quote_df_origin = subset(full_quote_df, !type %in% c("career-column", "news-and-views"))
full_quote_df_origin = unique(full_quote_df_origin)


# first we remove anything that is not a full name
# i.e. there must be a space
space_name_idx = grep(" ", full_quote_df_origin$est_speaker)
full_quote_df_origin = full_quote_df_origin[space_name_idx, ]

# now format the author name
full_quote_df_origin$author = format_author_fullname(full_quote_df_origin$est_speaker)

# format the output
col_ids = c("year", "type", "author", "file_id", "quote")
full_quote_df_origin = unique(full_quote_df_origin[,col_ids])

print("Total quotes with full name")
quotes_with_fullname = nrow(full_quote_df_origin)
print(quotes_with_fullname)

print("Total quotes with name origin prediction")
quotes_with_name_pred = nrow(unique(quote_name_origin_df))
print(quotes_with_name_pred)


table1_res = c(total_quotes, quotes_with_name_or_pronoun, quotes_with_gender, quotes_with_fullname, quotes_with_name_pred)
names(table1_res) = c("total_quotes", 
                      "quotes_with_name_or_pronoun", 
                      "quotes_with_gender_pred",
                      "quotes_with_fullname",
                      "quotes_with_name_pred")

```


### Name Origin prediction for citations
```{r citations_origin_pred_table2, echo=TRUE, warning=FALSE, message=F}

# first get all citations accessed
total_cite_name_df = cited_doi_df
journalist_idx = which(total_cite_name_df$file_id %in% news_df$file_id)
scientist_idx = which(!total_cite_name_df$file_id %in% news_df$file_id)
total_cite_name_df$corpus = ""
total_cite_name_df$corpus[journalist_idx] = "citation_journalist"
total_cite_name_df$corpus[scientist_idx] = "citation_scientist"

print("Number of Springer Citations")
total_cite_name_df = unique(total_cite_name_df)
tot_citations = data.frame(table(total_cite_name_df$corpus))
colnames(tot_citations) = c("type", "total_citations")
print(tot_citations)

# now get all citations accessed in springer
cite_name_df = cited_author_df

# seperate out citations from columns by journalists vs scientists
journalist_idx = which(cite_name_df$file_id %in% news_df$file_id)
scientist_idx = which(!cite_name_df$file_id %in% news_df$file_id)
cite_name_df$corpus = ""
cite_name_df$corpus[journalist_idx] = "citation_journalist"
cite_name_df$corpus[scientist_idx] = "citation_scientist"

print("Number of Springer Citations")
cite_name_df = unique(cite_name_df)
springer_citations = data.frame(table(cite_name_df$corpus))
colnames(springer_citations) = c("type", "total_springer_citations")
print(springer_citations)

# format the author names and add the correct corpus
cited_author_df_full = format_authors(cite_name_df, use_fullname=T)
cited_author_df_full = subset(cited_author_df_full, author != "")
journalist_idx = which(cited_author_df_full$file_id %in% news_df$file_id)
scientist_idx = which(!cited_author_df_full$file_id %in% news_df$file_id)
cited_author_df_full$corpus = ""
cited_author_df_full$corpus[journalist_idx] = "citation_journalist"
cited_author_df_full$corpus[scientist_idx] = "citation_scientist"

print("Number of Citations with full name")
cited_author_df_full = unique(cited_author_df_full)
named_citations = data.frame(table(cited_author_df_full$corpus, cited_author_df_full$author_pos))
named_citations = dcast(Var1 ~ Var2, data = named_citations, value.var="Freq")
colnames(named_citations) = c("type", "fullname_first_author_citations", "fullname_last_author_citations")
print(named_citations)

# filter to the authors with a name prediciton
pred_cite = subset(origin_cite_name_df, corpus %in% c("citation_journalist", "citation_scientist"))
print("Number of Citations with prediction")
pred_cite = unique(pred_cite)
pred_citations = data.frame(table(pred_cite$corpus, pred_cite$author_pos))
pred_citations = dcast(Var1 ~ Var2, data = pred_citations, value.var="Freq")
colnames(pred_citations) = c("type", 
                              "name_origin_pred_first_author_citations", 
                              "name_origin_pred_last_author_citations")
print(pred_citations)


table2_res = Reduce(merge, list(tot_citations, springer_citations, 
                                named_citations, pred_citations ))



```


### Gender + origin prediction for Springer
```{r springer_pred_table3, echo=TRUE, warning=FALSE, message=F}


springer_gender_df = springer_author_df
print("Number of Total Springer")
springer_gender_df = unique(springer_gender_df)
num_citations_springer = nrow(springer_gender_df)
print(num_citations_springer)

springer_gender_df = format_authors(springer_gender_df)
springer_gender_df = subset(springer_gender_df, author != "")
print("Number of Springer Full name")
springer_gender_df = unique(springer_gender_df)
num_names_springer = nrow(springer_gender_df)
print(num_names_springer)

res = get_author_gender(springer_gender_df)
springer_author_df_gender = res[[2]]
print("Number of Springer Full name")
springer_author_df_gender = subset(springer_author_df_gender, 
                                   !is.na(guessed_gender))
springer_author_df_gender = unique(springer_author_df_gender)
num_gender_springer = nrow(springer_author_df_gender)
print("Number of Springer Gender Prediction")
print(num_gender_springer)

# double check that it matches the pre-processed table we are working with
springer_pred_gender_df_pass = subset(springer_pred_gender_df, !is.na(guessed_gender))
springer_pred_gender_df_pass = unique(springer_pred_gender_df_pass)
num_names_springer_preprocessed = nrow(springer_pred_gender_df_pass)
print(num_names_springer_preprocessed)
print(num_names_springer_preprocessed == num_gender_springer)

pred_springer = subset(origin_cite_name_df, corpus  == "springer_articles")
print("Number of Springer with prediction")
pred_springer = unique(pred_springer)
pred_springer = nrow(pred_springer)
print(pred_springer)

table3_res = c(num_citations_springer, num_names_springer, num_gender_springer, pred_springer)
names(table3_res) = c("num_springer_articles", "num_springer_full_names", "num_springer_gender_pred", "num_springer_origin_pred")


```


### Gender + origin prediction for Nature
```{r nature_pred_table3, echo=TRUE, warning=FALSE, message=F}


nature_gender_df = nature_author_df
print("Number of Total Nature")
nature_gender_df = unique(nature_gender_df)
num_citations_nature = nrow(nature_gender_df)
print(num_citations_nature)

nature_gender_df = format_authors(nature_gender_df)
nature_gender_df = subset(nature_gender_df, author != "")
print("Number of Nature Full name")
nature_gender_df = unique(nature_gender_df)
num_names_nature = nrow(nature_gender_df)
print(num_names_nature)

res = get_author_gender(nature_gender_df)
nature_author_df_gender = res[[2]]
print("Number of nature Full name")
nature_author_df_gender = subset(nature_author_df_gender, 
                                   !is.na(guessed_gender))

num_gender_nature = unique(num_gender_nature)
num_gender_nature = nrow(nature_author_df_gender)
print("Number of nature Gender Prediction")
print(num_gender_nature)

# double check that it matches the pre-processed table we are working with
nature_pred_gender_df_pass = subset(nature_pred_gender_df, !is.na(est_gender))
num_names_nature_preprocessed = nrow(nature_pred_gender_df_pass)
print(num_names_nature_preprocessed)
print(num_names_nature_preprocessed == num_gender_nature)

origin_cite_name_df = unique(origin_cite_name_df)
pred_nature = subset(origin_cite_name_df, corpus  == "nature_articles")
print("Number of nature with prediction")
pred_nature = nrow(pred_nature)
print(pred_nature)

table4_res = c(num_citations_nature, num_names_nature, num_gender_nature, pred_nature)
names(table4_res) = c("num_nature_articles", "num_nature_full_names", "num_nature_gender_pred", "num_nature_origin_pred")

```

### Make Tables

```{r make_tables, echo=TRUE, warning=FALSE, message=F}

knitr::kable(data.frame(Frequency=table1_res), format = "pipe", 
             caption = "Breakdown of quotes at major processing steps")

knitr::kable(table2_res, format = "pipe", 
             caption = "Breakdown of citations at major processing steps")

knitr::kable(data.frame(Frequency=table3_res), format = "pipe", 
             caption = "Breakdown of all Springer articles at major processing steps")

knitr::kable(data.frame(Frequency=table4_res), format = "pipe", 
             caption = "Breakdown of all Nature articles at major processing steps")

```
