---
title: "Quote_Gender_Name_Origin"
author: "Natalie Davidson"
date: "5/3/2021"
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

require(data.table)
require(here)
require(ggplot2)
require(magick)
require(pdftools)
require(stringr)
require(ggrepel)
options(ggrepel.max.overlaps = Inf)

proj_dir = here()
source(file.path(proj_dir, "/utils/scraper_processing_utils.R"))
source(file.path(proj_dir, "/analysis_scripts/analysis_utils.R"))
source(file.path(proj_dir, "/utils/plotting_utils.R"))
source(file.path(proj_dir, "/process_doi_data/springer_scripts/springer_scrape_utils.R"))

RERUN_BOOTSTRAP = FALSE

name_origin_ids = c("African", "CelticEnglish", 
                    "EastAsian", "European", "Greek", "Hispanic", 
                    "Hebrew", "ArabTurkPers", "Nordic", "SouthAsian")

```


## Overview

This notebook generates a supplemental figure to highlight the interesection of gender and name origin.


The **data** it uses to build the plots are here:

This document compares mentions across predicted gender and name origin.

The names mentioned data file is: `./data/author_data/all_mentioned_fullname_pred.tsv`
The bg data file is: `./data/author_data/all_author_fullname_pred.tsv`


The three corpi are indexed by the `corpus` column:

1) `news_quotes`: __foreground__ est. name origin of Nature News quoted speaker 

2) `nature_last`: __background__ est. name origin of last author of Nature articles. 



The **setting + helper functions** to generate the plots are here:

1. plotting related functions: `/utils/plotting_utils.R`

2. reading + data processing related functions: `/utils/scraper_processing_utils.R` and `/analysis_scripts/analysis_utils.R`

3. nautre research article and springer specific data processing functions: `/process_doi_data/springer_scripts/springer_scrape_utils.R`




## Read in the data

```{r read_name_pred, echo=TRUE, warning=FALSE, message=F}

# read in raw quotes data for filtering
full_quote_df = NA
quote_files = list.files(file.path(proj_dir,"/data/scraped_data/", sep=""), full.names = T)
quote_files = grep("quote_table_raw_", quote_files, value=T)
for(quote_file in quote_files){
    
    quote_df = read_corenlp_quote_files(quote_file)
    quote_df$year = str_extract(quote_file, "[1-9][0-9]+") # curr_year
    quote_df$type = substring(basename(quote_file), 
                            22, nchar(basename(quote_file))-4)
    
    full_quote_df = rbind(full_quote_df, quote_df)
}
full_quote_df = full_quote_df[-1,]

# filter out articles with more than 25 quotes
num_quotes = table(full_quote_df$file_id)
too_many_quotes_idx = which(num_quotes > 25)
too_many_quotes_file_id = names(num_quotes)[too_many_quotes_idx]


# first read in the quote data
name_pred_file = file.path(proj_dir, 
                             "/data/author_data/all_speaker_fullname_pred.tsv")
name_info_file = file.path(proj_dir, 
                             "/data/author_data/all_speaker_fullname.tsv")

quote_name_df = read_name_origin(name_pred_file, name_info_file)
quote_name_df$name_origin[quote_name_df$name_origin == "Jewish"] = "Hebrew"
quote_name_df = subset(quote_name_df, !file_id %in% too_many_quotes_file_id)
quote_name_df = quote_name_df[which(quote_name_df$year <= 2020),]


# filter out career column and news-and-views
quote_name_df = subset(quote_name_df, !type %in% c("career-column", "news-and-views", "guardian"))
quote_name_df = unique(quote_name_df)


```

## Process Data


### get predicted gender estimate

```{r get_gender, echo=TRUE, warning=FALSE, message=F}

gender_info_file = file.path(proj_dir, 
                             "data/reference_data/genderize.tsv")
gender_update_file = file.path(proj_dir, 
                             "data/reference_data/genderize_update.tsv")

gender_df = data.frame(fread(gender_info_file))
gender_update_df = data.frame(fread(gender_update_file))
gender_df = rbind(gender_df, gender_update_df)

name_df_split = separate(data = quote_name_df, col = author, into = c("fore_name_simple", "author_lastname"), sep = " ")
name_df_split$fore_name_simple = tolower(name_df_split$fore_name_simple)
name_origin_gender_df = merge(name_df_split, gender_df)

name_origin_gender_df = name_origin_gender_df[which(!is.na(name_origin_gender_df$probability_male)),]
name_origin_gender_df$predicted_gender = "MALE"
name_origin_gender_df$predicted_gender[name_origin_gender_df$probability_male < 0.5] = "FEMALE"

```

### Format table for results

```{r bootstrap_est, echo=TRUE, warning=FALSE, message=F}

gender_origin = table(name_origin_gender_df$name_origin, name_origin_gender_df$predicted_gender)
ratio_male = gender_origin[,2] / (gender_origin[,1] + gender_origin[,2])
gender_origin = cbind(gender_origin, ratio_male)
gender_origin



knitr::kable(gender_origin, format = "pipe", 
             caption = "Quoted speaker gender by name origin")


```
