---
title: "Gender analysis"
author: "Natalie Davidson"
date: "1/12/2021"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

require(data.table)
require(here)
require(ggplot2)
require(caret)
require(stringr)


proj_dir = here()
source(paste(proj_dir, "/analysis_scripts/analysis_utils.R", sep=""))
source(paste(proj_dir, "/utils/plotting_utils.R", sep=""))
```

## Nature News Gender Disparities in Quotes

This document is a working analysis of the quotes extracted from Nature News content to see if there are differences in gender representation.
The data we will be working with are the following:

1) `./data/benchmark_data/benchmark_quote_table_raw.tsv` is the output after scraping a randomly selected set of 10 articles from 2010, 2015, or 2020 (`./nature_news_scraper/run_scrape_benchmark.sh`) then running it through coreNLP with additional processing (`./process_scraped_data/run_process_target_year.sh`)
2) `./data/scraped_data/quote_table_raw_20*.tsv` are the output after scraping all articles from a year between 2001 2020 (`./nature_news_scraper/run_scrape_benchmark.sh`) then running it through coreNLP with additional processing (`./process_scraped_data/run_process_target_year.sh`)


**All analysis shown below depends on the functions described in `/analysis_scripts/analysis_utils.R`**


## Check Benchmark Consistency

We would like to make sure that our benchmark data is representative of our complete dataset, so let's look at some quote stats to make sure.

### reading in the quote data

First we read in the benchmark data
```{r}

# get the project directory, everything is set relative to this
proj_dir = here()

# read in the benchmark quote table
raw_quote_file = paste(proj_dir, 
                    "/data/benchmark_data/benchmark_quote_table_raw.tsv", 
                    sep="")

bm_quote_df = read_corenlp_quote_files(raw_quote_file)

# add the year annotation
year_idx_file = paste(proj_dir, 
                "/data/benchmark_data/coreNLP_input/fileID_year.tsv", 
                sep="")
year_df = data.frame(fread(year_idx_file))
bm_quote_df = merge(year_df, bm_quote_df)


head(bm_quote_df)


```

Now we read in the full data for the same years.
```{r}

# read in the full year quote table for 2010, 2015, and 2020
full_quote_df = NA
quote_files = list.files(paste(proj_dir,"/data/scraped_data/", sep=""), full.names = T)
quote_files = grep("quote_table_raw_", quote_files, value=T)
for(quote_file in quote_files){
    
    quote_df = read_corenlp_quote_files(quote_file)
    quote_df$year = str_extract(quote_file, "[1-9][0-9]+") # curr_year

    full_quote_df = rbind(full_quote_df, quote_df)
}
full_quote_df = full_quote_df[-1,]

head(full_quote_df)


```

Now we join the tables together for comparison.
```{r}

full_quote_df$is_benchmark = "full_data"
bm_quote_df$is_benchmark = "benchmark"

bm_full_quote_df = rbind(full_quote_df, bm_quote_df)

```

### compare benchmark and non-benchmark data
```{r fig.align='center', echo=FALSE, warning=FALSE, message=F}


    ggplot(bm_full_quote_df, aes(x=is_benchmark, fill=est_gender)) +
        geom_bar(position="fill") + facet_grid(~year) + theme_bw() + 
        xlab("Year of Article") + ylab("Male Quote Percentage") +
        ggtitle("Male Quote Percentage Over Time") + 
        scale_fill_brewer(palette="Set2")


```

### compare proportions over all years

Now lets look at all years quote information and plot the trend over time
```{r fig.align='center', echo=FALSE, warning=FALSE, message=F}
    
    # read in the additional years and append to `full_suote_df`
    # for(curr_year in c(2016:2019)){
    #     quote_file = paste(proj_dir, 
    #                     "/data/scraped_data/quote_table_raw_", curr_year, ".tsv", 
    #                     sep="")
    #     quote_df = read_corenlp_quote_files(quote_file)
    #     quote_df$year = curr_year
    #     quote_df$is_benchmark = "full_data"
    # 
    #     full_quote_df = rbind(full_quote_df, quote_df)
    # }

    ggplot(full_quote_df, aes(x=as.factor(year), fill=est_gender)) +
        geom_bar(position="fill") + theme_bw() + 
        xlab("Year of Article") + ylab("Male Quote Percentage") +
        ggtitle("Male Quote Percentage Over Time") + 
        scale_fill_brewer(palette="Set2")
    
     ggplot(subset(full_quote_df, est_gender %in% c("MALE", "FEMALE")), aes(x=as.factor(year), fill=est_gender)) +
        geom_bar(position="fill") + theme_bw() + 
        xlab("Year of Article") + ylab("Male Quote Percentage") +
        ggtitle("Male Quote Percentage Over Time") + 
        scale_fill_brewer(palette="Set2")


```