---
title: "location_analysis"
author: "Natalie Davidson"
date: "1/22/2021"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

require(data.table)
require(here)
require(ggplot2)
require(ggrepel)
require(ggpubr)
require(caret)

proj_dir = here()
source(paste(proj_dir, "/analysis_scripts/analysis_utils.R", sep=""))
source(paste(proj_dir, "/utils/plotting_utils.R", sep=""))
```

## Nature News Location Bias

This document is a working analysis of the quotes extracted from Nature News content to see if there are differences in gender representation.
The data we will be working with are the following:

1) `./data/benchmark_data/benchmark_quote_table_raw.tsv` is the output after scraping a randomly selected set of 10 articles from 2010, 2015, or 2020 (`./nature_news_scraper/run_scrape_benchmark.sh`) then running it through coreNLP with additional processing (`./process_scraped_data/run_process_target_year.sh`)
2) `./data/scraped_data/quote_table_raw_20*.tsv` are the output after scraping all articles from a year between 2001 2020 (`./nature_news_scraper/run_scrape_benchmark.sh`) then running it through coreNLP with additional processing (`./process_scraped_data/run_process_target_year.sh`)


**All analysis shown below depends on the functions described in `/analysis_scripts/analysis_utils.R`**


## Check Benchmark Consistency

We would like to make sure that our benchmark data is representative of our complete dataset, so let's look at some quote stats to make sure.

### reading in the quote data

First we read in the benchmark data
```{r}

# get the project directory, everything is set relative to this
proj_dir = here()

# read in the benchmark quote table
bm_loc_file = paste(proj_dir, 
                    "/data/benchmark_data/benchmark_location_table_raw.tsv", 
                    sep="")

bm_loc_df = read_corenlp_location_files(bm_loc_file)

# add the year annotation
year_idx_file = paste(proj_dir, 
                "/data/benchmark_data/coreNLP_input/fileID_year.tsv", 
                sep="")
year_df = data.frame(fread(year_idx_file))
bm_loc_df = merge(year_df, bm_loc_df)


head(bm_loc_df)


```

Now we read in the full data for the same years.
```{r}

# read in the full year quote table for 2010, 2015, and 2020
full_loc_df = NA
for(curr_year in c(2010, 2015, 2020)){
    loc_file = paste(proj_dir, 
                    "/data/scraped_data/location_table_raw_", curr_year, ".tsv", 
                    sep="")
    loc_df = read_corenlp_location_files(loc_file)
    loc_df$year = curr_year

    full_loc_df = rbind(full_loc_df, loc_df)
}
full_loc_df = full_loc_df[-1,]

head(full_loc_df)


```

Now we join the tables together for comparison.
```{r}

full_loc_df$is_benchmark = "full_data"
bm_loc_df$is_benchmark = "benchmark"

bm_full_loc_df = rbind(full_loc_df, bm_loc_df)
bm_full_loc_df = subset(bm_full_loc_df, est_un_region != "" & 
                                        est_un_subregion != "" &
                                        est_un_region != "NO_EST" & 
                                        est_un_subregion != "NO_EST")

```

### compare benchmark and non-benchmark data

Now lets check how well the benchmark proportions match the full dataset proportions.

## Lets look at the UN region level
```{r fig.align='center', fig.width = 15, fig.height = 10, echo=FALSE, warning=FALSE, message=F}

    
    num_articles = unique(bm_full_loc_df[,c("is_benchmark", "year", 
                                            "file_id")])
    num_articles = num_articles %>%
                    group_by(is_benchmark, year) %>% 
                    summarize(n())
    colnames(num_articles) = c("is_benchmark", "year",
                               "num_total_articles")

    # function for aggregating over the region of interest
    get_aggr_region <- function(bm_full_loc_df, region_col_id, num_articles){
        
        region_df = unique(bm_full_loc_df[,c("is_benchmark", "year", 
                                            region_col_id, "file_id")])
        colnames(region_df)[3] = "est_region"
        region_df = region_df %>%
                        group_by(is_benchmark, year, est_region) %>% 
                        summarize(n())
        colnames(region_df) = c("is_benchmark", "year",
                                   "est_region", "num_articles")
        
        region_df = merge(num_articles, region_df)
        region_df$prop_mentioned = region_df$num_articles / region_df$num_total_articles
        
        # plotting labels
        region_df$label = region_df$est_region
        region_df$label[region_df$is_benchmark == "benchmark"] = ""
        
        colnames(region_df)[4] = region_col_id

        
        return(region_df)

    }
    
    un_region_df = get_aggr_region(bm_full_loc_df, region_col_id="est_un_region", 
                                   num_articles)
    ggplot(un_region_df, aes(x=is_benchmark, y=prop_mentioned, 
                             color = est_un_region, fill=est_un_region,
                             label = label)) +
        geom_point() + geom_line(aes(group=est_un_region)) + 
        geom_text_repel() + facet_grid(~year) + theme_bw() + 
        xlab("Year of Article") + 
        ylab("Fraction Articles with mention of UN Region ") +
        ggtitle("Proportion of articles with at least 1 mention of a UN Region") + 
        scale_fill_brewer(palette="Set2") + ylim(c(0, 1.1)) +
        theme(legend.position = "none")
    
    

```

## Lets look at the UN Subregion level
```{r fig.align='center', fig.width = 15, fig.height = 10, echo=FALSE, warning=FALSE, message=F}

    
    un_subregion_df = get_aggr_region(bm_full_loc_df, region_col_id="est_un_subregion", 
                                   num_articles)
    ggplot(un_subregion_df, aes(x=is_benchmark, y=prop_mentioned, 
                             color = est_un_subregion, fill=est_un_subregion,
                             label = label)) +
        geom_point() + geom_line(aes(group=est_un_subregion)) + 
        geom_text_repel() + facet_grid(~year) + theme_bw() + 
        xlab("Year of Article") + 
        ylab("Fraction Articles with mention of UN Subregion ") +
        ggtitle("Proportion of articles with at least 1 mention of a UN Subregion") + 
        scale_fill_brewer(palette="Set2") + ylim(c(0, 1.1)) +
        theme(legend.position = "none")
    
    prop_region = subset(un_subregion_df, prop_mentioned < 0.31 )
    ggplot(subset(un_subregion_df, est_un_subregion %in% prop_region$est_un_subregion),
           aes(x=is_benchmark, y=prop_mentioned, 
                 color = est_un_subregion, fill=est_un_subregion,
                 label = label)) +
        geom_point() + geom_line(aes(group=est_un_subregion)) + 
        geom_text_repel() + facet_grid(~year) + theme_bw() + 
        xlab("Year of Article") + 
        ylab("Fraction Articles with mention of UN Subregion ") +
        ggtitle("Proportion of articles with at least 1 mention of a UN Subregion
                subsetted to having a proportion 0.3 or lower") + 
        scale_fill_brewer(palette="Set2") +
        theme(legend.position = "none")


```

## Now Finally the Country Code level, except divided by subregion
```{r fig.align='center', fig.width = 15, fig.height = 10, echo=FALSE, warning=FALSE, message=F}

    
    un_cc_df = get_aggr_region(bm_full_loc_df, region_col_id="est_country_code", 
                                   num_articles)
    subregion_ref = unique(bm_full_loc_df[,c("est_country_code", "est_un_subregion")])
    
    bg_obs = subset(un_cc_df, is_benchmark == "benchmark" & 
                           num_articles > 0)
    un_cc_df = subset(un_cc_df, est_country_code %in% bg_obs$est_country_code)

    un_cc_df = merge(un_cc_df, subregion_ref)

    all_gg = list()
    for(curr_subregion in unique(un_cc_df$est_un_subregion)){
        
        plot_df = subset(un_cc_df, est_un_subregion == curr_subregion)
        gg = ggplot(plot_df, aes(x=is_benchmark, y=prop_mentioned, 
                             color = est_country_code, fill=est_country_code,
                             label = label)) +
            geom_point() + geom_line(aes(group=est_country_code)) +
            geom_text_repel() + 
            facet_grid(~year) + theme_bw() + 
            xlab("Year of Article") + 
            ylab("Fraction Articles with mention of Country") +
            ggtitle(paste(curr_subregion, " --  Proportion of articles with at least 1 mention of a Country on BG")) + 
            theme(legend.position = "none")
        all_gg[[curr_subregion]] = gg
    }
    ggarrange(plotlist=all_gg, ncol = 2)
    

```