---
title: "name_origin_quotes_with_bg"
author: "Natalie Davidson"
date: "4/28/2021"
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

require(data.table)
require(here)
require(ggplot2)
require(ggrepel)
require(stringr)


proj_dir = here()
source(file.path(proj_dir, "/analysis_scripts/analysis_utils.R"))
source(file.path(proj_dir, "/utils/plotting_utils.R"))
source(file.path(proj_dir, "/utils/scraper_processing_utils.R"))

```

## Data Description

This document compares one "foreground" datasets (estimated name origin of authors quoted in nature news articles) and compares it to two possible "background" datasets (random sampling of 36K Springer articles, and all nature articles)

The quote data file is: `./data/author_data/all_speaker_fullname_pred.tsv`
The bg data file is: `./data/author_data/all_author_fullname_pred.tsv`

The three corpi are indexed by the `corpus` column:

1) `news_quotes`: __foreground__ est. name origin of Nature News quoted speaker 

2) `nature_articles`: __background__ est. name origin of last author of Nature articles. 

3) `springer_articles`: __background__ est. name origin of last author of a random subset of Springer articles. 


## Foreground Location Breakdown


Read in the name data.

```{r}
# get the project directory, everything is set relative to this
proj_dir = here()

#' public method that reads in the name origine prediction 
#' files and formats it into a dataframe
#' 
#' @param name_pred_file, prediction results from Wiki2019-LSTM
#' @param name_info_file, file used as input to Wiki2019-LSTM
#' @return dataframe of name origin predictions with annotation info
read_name_origin <- function(name_pred_file, name_info_file){
        
    # read in the name data
    name_pred_df = data.frame(fread(name_pred_file))
    name_info_df = data.frame(fread(name_info_file))
    
    # format the prediction table
    colnames(name_pred_df)[1] = "author"
    name_origin_vec = colnames(name_pred_df)[2:ncol(name_pred_df)]
    name_origin = apply(name_pred_df[,2:ncol(name_pred_df)], 1, 
                        function(x) name_origin_vec[which(x == max(x))])
    name_pred_df$name_origin = name_origin
    
    name_df = merge(name_info_df, name_pred_df[,c("author", "name_origin")])
    
    # remove any names that may have got through that are not real names
    name_df = name_df[grep("collab|group", tolower(name_df$author), invert=T), ]
    name_df = unique(name_df)
    
    
    return(name_df)
}

# first read in the quote data
name_pred_file = file.path(proj_dir, 
                             "/data/author_data/all_speaker_fullname_pred.tsv")
name_info_file = file.path(proj_dir, 
                             "/data/author_data/all_speaker_fullname.tsv")

quote_name_df = read_name_origin(name_pred_file, name_info_file)


# now read in the BG data
name_pred_file = file.path(proj_dir, 
                         "/data/author_data/all_author_fullname_pred.tsv")
name_info_file = file.path(proj_dir, 
                         "/data/author_data/all_author_fullname.tsv")
cite_name_df = read_name_origin(name_pred_file, name_info_file)

# we will only use last authors in the citations
cite_name_df = subset(cite_name_df, author_pos == "last")

# now we want to join these two datasets together
# we assume a quote is comparable to a publication
# so we will have a quote set as a doi
quote_name_df$doi = quote_name_df$quote
quote_name_df$corpus = "naturenews_quotes"


col_ids = c("author", "year", "name_origin", "corpus", "doi")
name_df = rbind(cite_name_df[,col_ids], quote_name_df[,col_ids])

head(name_df)

```

### QC

From each article type, lets compare the total number of articles considered in this analysis.


```{r qc_analysis, fig.align='center', echo=FALSE, warning=FALSE, message=F}

quote_name_df_type = unique(subset(quote_name_df, select=c("type", "file_id", "year")))
ggplot(quote_name_df_type, aes(x=as.numeric(year), fill=type)) +
    geom_bar() + theme_bw() + 
    xlab("Year of Article") + ylab("Number of Total Articles") +
    ggtitle("Total Number of Articles with a Quoted Speaker per Type") + 
    scale_fill_brewer(palette="Set2")


citation_total = unique(subset(name_df, corpus == "naturenews_citations", select=c(doi, year)) )
tot_prop_citation = citation_total %>% 
                group_by(year) %>% 
                summarise(n()) 
tot_prop_citation$corpus = "naturenews_citations"

quote_total = unique(subset(name_df, corpus == "naturenews_quotes", select=c(doi, year)) )
tot_prop_quote = quote_total %>% 
                group_by(year) %>% 
                summarise(n()) 
tot_prop_quote$corpus = "naturenews_quotes"

springer_total = unique(subset(name_df, corpus == "springer_articles", select=c(doi, year)) )
tot_prop_springer = springer_total %>% 
                group_by(year) %>% 
                summarise(n()) 
tot_prop_springer$corpus = "springer_articles"

nature_total = unique(subset(name_df, corpus == "nature_articles", select=c(doi, year)) )
tot_prop_nature = nature_total %>% 
                group_by(year) %>% 
                summarise(n()) 
tot_prop_nature$corpus = "nature_articles"

num_art_tot = Reduce(rbind, list(tot_prop_citation, 
                                 tot_prop_quote,
                                 tot_prop_springer, 
                                 tot_prop_nature))
num_art_tot = data.frame(num_art_tot)
colnames(num_art_tot)[2] = "tot_articles"
 
ggplot(num_art_tot, aes(x=as.numeric(year), y=tot_articles,
                              fill=corpus, color=corpus)) +
    geom_point() + geom_line() + theme_bw() + 
    xlab("Year of Article") + ylab("Number of Total Articles") +
    ggtitle("Total number of Articles per Corpus") + 
    scale_fill_brewer(palette="Set2")



```

### Name Origin Comparison

For each corpus, lets compare the number of articles with an estimated name origin for the first and last authors.


```{r compare_origins_each_corpus, fig.align='center', fig.width = 10, fig.height = 10, echo=FALSE, warning=FALSE, message=F}


# make the df into proportions for each name origins
get_subboot <- function(origin_id, curr_corpus, in_df, bootstrap_col_id="doi"){
    bootstrap_res = compute_bootstrap_location(subset(in_df, 
                                                      corpus==curr_corpus), 
                                              year_col_id = "year", 
                                              article_col_id = bootstrap_col_id, 
                                              country_col_id = "name_origin",
                                              country_agg = origin_id, 
                                              conf_int = 0.95)
    bootstrap_res$name_origin = origin_id
    return(bootstrap_res)

}

# get the bootstrapped CI
citation_origin_df = NA
for(curr_origin in unique(name_df$name_origin)){
    print(curr_origin)
    res = get_subboot(curr_origin, 
                      curr_corpus="naturenews_citations",
                      name_df)
    citation_origin_df = rbind(citation_origin_df, res)
}
citation_origin_df = citation_origin_df[-1,]

quote_origin_df = NA
for(curr_origin in unique(name_df$name_origin)){
    print(curr_origin)
    res = get_subboot(curr_origin, 
                      curr_corpus="naturenews_quotes",
                      name_df)
    quote_origin_df = rbind(quote_origin_df, res)
}
quote_origin_df = quote_origin_df[-1,]


springer_origin_df = NA
for(curr_origin in unique(name_df$name_origin)){
    print(curr_origin)
    res = get_subboot(curr_origin, 
                      curr_corpus="springer_articles",
                      name_df)
    springer_origin_df = rbind(springer_origin_df, res)
}
springer_origin_df = springer_origin_df[-1,]

nature_origin_df = NA
for(curr_origin in unique(name_df$name_origin)){
    print(curr_origin)
    res = get_subboot(curr_origin, 
                      curr_corpus="nature_articles",
                      name_df)
    nature_origin_df = rbind(nature_origin_df, res)
}
nature_origin_df = nature_origin_df[-1,]

# format them for plotting

citation_origin_df$label = ""
citation_origin_df$label[citation_origin_df$year == 2020] = 
    citation_origin_df$name_origin[citation_origin_df$year == 2020]
    
quote_origin_df$label = ""
quote_origin_df$label[quote_origin_df$year == 2020] = 
    quote_origin_df$name_origin[quote_origin_df$year == 2020]

springer_origin_df$label = ""
springer_origin_df$label[springer_origin_df$year == 2020] = 
    springer_origin_df$name_origin[springer_origin_df$year == 2020]

nature_origin_df$label = ""
nature_origin_df$label[nature_origin_df$year == 2020] = 
    nature_origin_df$name_origin[nature_origin_df$year == 2020]

ggplot(citation_origin_df, aes(x=as.numeric(year), y=mean,
                          ymin=bottom_CI, ymax=top_CI,
                          fill=name_origin, label=label)) +
    geom_point() + geom_ribbon(alpha=0.5) + geom_line(alpha=0.5) +
    theme_bw() + geom_text_repel() + xlim(c(2005,2021))  +
    xlab("Year of Article") + ylab("Proportion articles from specific name origins") +
    ggtitle("Estimated Proportion of Nature News Cited articles of Last Author name origin") + 
    scale_fill_brewer(palette="Set3") +
    theme(legend.position = "none")


ggplot(quote_origin_df, aes(x=as.numeric(year), y=mean,
                          ymin=bottom_CI, ymax=top_CI,
                          fill=name_origin, label=label)) +
    geom_point() + geom_ribbon(alpha=0.5) + geom_line(alpha=0.5) +
    theme_bw() + geom_text_repel() + xlim(c(2005,2021))  +
    xlab("Year of Article") + ylab("Proportion articles from specific name origins") +
    ggtitle("Estimated Proportion of Nature News Quotes name origin") + 
    scale_fill_brewer(palette="Set3") +
    theme(legend.position = "none")


ggplot(springer_origin_df, aes(x=as.numeric(year), y=mean,
                          ymin=bottom_CI, ymax=top_CI,
                          fill=name_origin, label=label)) +
    geom_point() + geom_ribbon(alpha=0.5) + geom_line(alpha=0.5) +
    theme_bw() + geom_text_repel() + xlim(c(2005,2021))  +
    xlab("Year of Article") + ylab("Proportion articles from specific name origins") +
    ggtitle("Estimated Proportion of Springer articles of Last Author name origin") + 
    scale_fill_brewer(palette="Set3") +
    theme(legend.position = "none")

ggplot(nature_origin_df, aes(x=as.numeric(year), y=mean,
                          ymin=bottom_CI, ymax=top_CI,
                          fill=name_origin, label=label)) +
    geom_point() + geom_ribbon(alpha=0.5) + geom_line(alpha=0.5) +
    theme_bw() + geom_text_repel() + xlim(c(2005,2021))  +
    xlab("Year of Article") + ylab("Proportion articles from specific name origins") +
    ggtitle("Estimated Proportion of Nature articles of Last Author name origin") + 
    scale_fill_brewer(palette="Set3") +
    theme(legend.position = "none")



```


Lets look at individual name origins

```{r individual_name_origins, out.width="50%", echo=FALSE, warning=FALSE, message=F}


citation_origin_df$corpus = "naturenews_citations"
quote_origin_df$corpus = "naturenews_quotes"
springer_origin_df$corpus = "springer_articles"
nature_origin_df$corpus = "nature_articles"
foreground_df = Reduce(rbind, list(quote_origin_df,
                                   nature_origin_df))

for(curr_origin in unique(foreground_df$name_origin)){
    
    gg = ggplot(subset(foreground_df, name_origin == curr_origin), 
       aes(x=as.numeric(year), y=mean,
                          ymin=bottom_CI, ymax=top_CI,
                          fill=corpus)) +
    geom_point() + geom_ribbon(alpha=0.5) + geom_line(alpha=0.5) +
    theme_bw() + 
    xlab("Year of Article") + ylab("Percentage articles with last author name origin") +
    ggtitle(paste("Percentage articles with", curr_origin, "name origin")) + 
    scale_fill_brewer(palette="Set2")
    
    print(gg)
    
}

foreground_df = Reduce(rbind, list(quote_origin_df,
                                   springer_origin_df))
for(curr_origin in unique(foreground_df$name_origin)){
    
    gg = ggplot(subset(foreground_df, name_origin == curr_origin), 
       aes(x=as.numeric(year), y=mean,
                          ymin=bottom_CI, ymax=top_CI,
                          fill=corpus)) +
    geom_point() + geom_ribbon(alpha=0.5) + geom_line(alpha=0.5) +
    theme_bw() + 
    xlab("Year of Article") + ylab("Percentage articles with last author name origin") +
    ggtitle(paste("Percentage articles with", curr_origin, "name origin")) + 
    scale_fill_brewer(palette="Set2")
    
    print(gg)
    
}

```