---
title: "location_with_bg_analysis"
author: "Natalie Davidson"
date: "3/2/2021"
output: rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

require(data.table)
require(here)
require(ggplot2)
require(ggrepel)
require(stringr)
library(pheatmap)


proj_dir = here()
source(file.path(proj_dir, "/analysis_scripts/analysis_utils.R"))
source(file.path(proj_dir, "/utils/plotting_utils.R"))
source(file.path(proj_dir, "/utils/scraper_processing_utils.R"))

```

## Data Description

This document compares two "foreground" datasets (locations mentioned  and locations of authors cited in nature news articles) and compares it to two possible "background" datasets (random sampling of 36K Springer articles, and all nature articles)

The source data file is: `./data/author_data/all_author_country.tsv`

The four corpi are indexed by the `corpus` column:

1) `nature_news`: __foreground__ country of a location mentioned in any Nature News article

2) `news_citation`: __foreground__ country of Nature News cited authors affiliation. 

3) `nature_articles`: __background__ country of author affiliation from Nature articles. 

4) `springer`: __background__ country of author affiliation from random subset of Springer articles. 

The `num_entries` column denotes the number of articles with at least ONE author from a particular country
The `address.country_code` column denotes the UN 2-digit country code



A reference data file is:  `./data/author_data/total_num_articles_per_corpus.tsv`
This file contains the number of articles per corpus for reference.

## Foreground Location Breakdown


Read in the country data from all sources.
```{r}
# get the project directory, everything is set relative to this
proj_dir = here()

# read in the cited author data
country_file = file.path(proj_dir, "/data/author_data/all_author_country.tsv")
country_df = data.frame(fread(country_file))

un_info = get_country_info()
country_df = merge(country_df, un_info)


```

### compare cited vs mentioned regions over all years

From the Nature News corpus, lets compare the countries of locations mentioned in Nature news articles against the countries of cited authors.

Here lets first look at the total number of articles considered (number of nature news articles per year, and the number of articles cited by Nature News and indexed by Springer)


```{r foreground_analysis, fig.align='center', echo=FALSE, warning=FALSE, message=F}

# now lets add a QC plot 
mention_total = unique(subset(country_df, corpus == "naturenews_mentions", select=c(file_id, year)) )
tot_prop_mention = mention_total %>% 
                group_by(year) %>% 
                summarise(n()) 
tot_prop_mention$corpus = "naturenews_mentions"

citation_total = unique(subset(country_df, corpus == "naturenews_citations", select=c(file_id, year)) )
tot_prop_citation = citation_total %>% 
                group_by(year) %>% 
                summarise(n()) 
tot_prop_citation$corpus = "naturenews_citations"

num_art_foreground = rbind(tot_prop_mention, tot_prop_citation)
num_art_foreground = data.frame(num_art_foreground)
colnames(num_art_foreground)[2] = "tot_articles"
 
ggplot(num_art_foreground, aes(x=as.numeric(year), y=tot_articles,
                              fill=corpus, color=corpus)) +
    geom_point() + geom_line() + theme_bw() + 
    xlab("Year of Article") + ylab("Number of Total Articles") +
    ggtitle("Total number of Articles per Foreground Corpus") + 
    scale_fill_brewer(palette="Set2")

```

Let's first compare different UN subregions to one another in the two cohorts.

```{r compare_cite_v_mention_un_subregion, fig.align='center', fig.width = 10, fig.height = 10, echo=FALSE, warning=FALSE, message=F}

top_un_subregions = c("Northern America", "Western Europe", "Northern Europe", "Eastern Asia")
top_un_subregions = subset(country_df, un_subregion %in% top_un_subregions)
top_un_subregions = unique(top_un_subregions$country)
top_un_subregions = top_un_subregions[c(33,13, 6,7,12, 4, 5, 29, 20)]
top_un_subregions = c(top_un_subregions, "India")

# make the df into proportions for eachcountry
get_subboot <- function(country_id, curr_corpus, in_df){
    bootstrap_res = compute_bootstrap_location(subset(in_df, corpus==curr_corpus), 
                                              year_col_id = "year", 
                                              article_col_id = "file_id", 
                                              country_col_id = "country",
                                              country_agg = country_id, 
                                              conf_int = 0.95)
    bootstrap_res$country = country_id
    return(bootstrap_res)

}


mention_country_df = NA
for(curr_country in top_un_subregions){
    print(curr_country)
    res = get_subboot(curr_country, 
                      curr_corpus="naturenews_mentions", 
                      country_df)
    mention_country_df = rbind(mention_country_df, res)
}
mention_country_df = mention_country_df[-1,]


citation_country_df = NA
for(curr_country in top_un_subregions){
    print(curr_country)
    res = get_subboot(curr_country, 
                      curr_corpus="naturenews_citations", 
                      country_df)
    citation_country_df = rbind(citation_country_df, res)
}
citation_country_df = citation_country_df[-1,]


mention_country_df$label = ""
mention_country_df$label[mention_country_df$year == 2020] = 
    mention_country_df$country[mention_country_df$year == 2020]
    
citation_country_df$label = ""
citation_country_df$label[citation_country_df$year == 2020] = 
    citation_country_df$country[citation_country_df$year == 2020]

ggplot(mention_country_df, aes(x=as.numeric(year), y=mean,
                          ymin=bottom_CI, ymax=top_CI,
                          fill=country, label=label)) +
    geom_point() + geom_ribbon(alpha=0.5) + geom_line(alpha=0.5) +
    theme_bw() + geom_text_repel() + xlim(c(2005,2021))  +
    xlab("Year of Article") + ylab("Comparison of Countries in top UN Subregion mention rates") +
    ggtitle("Proportion articles with mention of a Top UN Subregion locations") + 
    scale_fill_brewer(palette="Set3") +
    theme(legend.position = "none")

ggplot(citation_country_df, aes(x=as.numeric(year), y=mean,
                          ymin=bottom_CI, ymax=top_CI,
                          fill=country, label=label)) +
    geom_point() + geom_ribbon(alpha=0.5) + geom_line(alpha=0.5) +
    theme_bw() + geom_text_repel()  + xlim(c(2005,2021)) + 
    xlab("Year of Article") + ylab("Comparison of Countries in top UN Subregion citation rates") +
    ggtitle("Proportion articles with atleast one author affiliation in a UN Subregion") + 
    scale_fill_brewer(palette="Set3") +
    theme(legend.position = "none")




```



Now lets look at the proportion of articles with atleast 1 country mention or atleast 1 authors' affiliate country cited by Nature News. 

We first look at individual countries.

```{r four_countries_mention_or_citation, out.width="50%", echo=FALSE, warning=FALSE, message=F}


mention_country_df$corpus = "naturenews_mentions"
citation_country_df$corpus = "naturenews_citations"
foreground_df = rbind(mention_country_df, citation_country_df)


ggplot(subset(foreground_df, country == "United States"), 
       aes(x=as.numeric(year), y=mean,
                          ymin=bottom_CI, ymax=top_CI,
                          fill=corpus)) +
    geom_point() + geom_ribbon(alpha=0.5) + geom_line(alpha=0.5) +
    theme_bw() + 
    xlab("Year of Article") + ylab("Percentage articles with mention/citation of US locations") +
    ggtitle("Percentage articles with mention/citation of US locations") + 
    scale_fill_brewer(palette="Set2")

ggplot(subset(foreground_df, country == "United Kingdom"), 
       aes(x=as.numeric(year), y=mean,
                          ymin=bottom_CI, ymax=top_CI,
                          fill=corpus)) +
    geom_point() + geom_ribbon(alpha=0.5) + geom_line(alpha=0.5) +
    theme_bw() + 
    xlab("Year of Article") + ylab("Percentage articles with mention/citation of UK locations") +
    ggtitle("Percentage articles with mention/citation of UK locations") + 
    scale_fill_brewer(palette="Set2")

ggplot(subset(foreground_df, country == "People's Republic of China"), 
       aes(x=as.numeric(year), y=mean,
                          ymin=bottom_CI, ymax=top_CI,
                          fill=corpus)) +
    geom_point() + geom_ribbon(alpha=0.5) + geom_line(alpha=0.5) +
    theme_bw() + 
    xlab("Year of Article") + ylab("Percentage articles with mention/citation of Chinese locations") +
    ggtitle("Percentage articles with mention/citation of Chinese locations") + 
    scale_fill_brewer(palette="Set2")


ggplot(subset(foreground_df, country == "India"), 
       aes(x=as.numeric(year), y=mean,
                          ymin=bottom_CI, ymax=top_CI,
                          fill=corpus)) +
    geom_point() + geom_ribbon(alpha=0.5) + geom_line(alpha=0.5) +
    theme_bw() + 
    xlab("Year of Article") + ylab("Percentage articles with mention/citation of Indian locations") +
    ggtitle("Percentage articles with mention/citation of Indian locations") + 
    scale_fill_brewer(palette="Set2")



```


Now lets take the mention proportion - citation proportion for each country. 
This will help us understand if some countries are studied more or publish more, or its equal.

```{r mention_minus_citation, fig.align='center', fig.width = 15, fig.height = 10, echo=FALSE, warning=FALSE, message=F}


mention_mean_df = mention_country_df
citation_mean_df = citation_country_df

colnames(mention_mean_df)[4] = "mention_prop"
colnames(citation_mean_df)[4] = "citation_prop"
foreground_df = merge(mention_mean_df[,c("year", "country", "mention_prop")], 
                      citation_mean_df[,c("year", "country", "citation_prop")])

foreground_df$diff = foreground_df$mention_prop - foreground_df$citation_prop

foreground_matr = foreground_df %>%
                    dcast(country ~ year, value.var="diff")
row.names(foreground_matr) = foreground_matr$country
foreground_matr = foreground_matr[,-1]


top_diff_country_idx = which(apply(foreground_matr, 1, mean) > 0.02)
sub_matr = foreground_matr[top_diff_country_idx,]
max_val = max(abs(sub_matr))
breaks = c(seq(-1*max_val, max_val, by = 0.01))
color_pmap <- colorRampPalette(c("yellow", "white", "blue"))(length(breaks))
pheatmap(sub_matr, cluster_rows = T, 
         cluster_cols = F, display_numbers = T, 
         main = "Mentioned more than cited
                (Mention - Citation) Proportions, 
                Only showning Mean(Mention - Citation) > 0.02 countries",
         color = color_pmap, breaks = breaks)

bottom_diff_country_idx = which(apply(foreground_matr, 1, mean) < -0.01)
sub_matr = foreground_matr[bottom_diff_country_idx,]
max_val = max(abs(sub_matr))
breaks = c(seq(-1*max_val, max_val, by = 0.01))
color_pmap <- colorRampPalette(c("yellow", "white", "blue"))(length(breaks))
pheatmap(sub_matr, cluster_rows = T, 
         cluster_cols = F, display_numbers = T, 
         main = "Cited more than Mentioned
                (Mention - Citation) Proportions, 
                Only showning Mean(Mention - Citation) < 0 countries",
         color = color_pmap, breaks = breaks)

```


## Background location Breakdown


Now aggregate the background data: all Springer articles and all Nature articles.


Here lets first look at the total number of articles considered (number of nature articles per year, and the number of Springer articles)


```{r qc_num_article_background, out.width="50%", echo=FALSE, warning=FALSE, message=F}

# nature bg
nature_total = unique(subset(country_df, corpus == "nature_articles", select=c(file_id, year)) )
tot_prop_naturebg = nature_total %>% 
                group_by(year) %>% 
                summarise(n()) 
tot_prop_naturebg$corpus = "nature_articles"
colnames(tot_prop_naturebg)[2] = "num_entries"

# springer bg
springer_total = unique(subset(country_df, corpus == "springer_articles", select=c(file_id, year)) )
tot_prop_springerbg = springer_total %>% 
                group_by(year) %>% 
                summarise(n()) 
tot_prop_springerbg$corpus = "springer_articles"
colnames(tot_prop_springerbg)[2] = "num_entries"
tot_prop_springerbg$corpus = "springer_articles"

num_art_background = rbind(tot_prop_naturebg, 
                           tot_prop_springerbg[,colnames(tot_prop_naturebg)])
 
ggplot(num_art_background, aes(x=as.numeric(year), y=num_entries,
                              fill=corpus, color=corpus)) +
    geom_point() + geom_line() + theme_bw() + 
    xlab("Year of Article") + ylab("Number of Total Articles") +
    ggtitle("Total number of Articles per Background Corpus") + 
    scale_fill_brewer(palette="Set2") +
    scale_y_continuous(trans='log10')

```

So Springer has many more articles than Nature. 
Let's look comparatively at different countries to check their frequencies.
We see that Nature is very biased towards US/UK in comparison to springer.
I believe springer has non-english journals, but needs to be checked.


```{r springer_v_nature, out.width="50%", echo=FALSE, warning=FALSE, message=F}

nature_country_df = NA
for(curr_country in top_un_subregions){
    print(curr_country)
    res = get_subboot(curr_country, 
                      curr_corpus="nature_articles", 
                      country_df)
    nature_country_df = rbind(nature_country_df, res)
}
nature_country_df = nature_country_df[-1,]

nature_country_df$label = ""
nature_country_df$label[nature_country_df$year == 2020] = 
    nature_country_df$country[nature_country_df$year == 2020]
nature_country_df$corpus = "nature_articles"

# springer bg
springer_country_df = NA
for(curr_country in top_un_subregions){
    print(curr_country)
    res = get_subboot(curr_country, 
                      curr_corpus="springer_articles", 
                      country_df)
    springer_country_df = rbind(springer_country_df, res)
}
springer_country_df = springer_country_df[-1,]

springer_country_df$corpus = "springer_articles"
springer_country_df$label = ""
springer_country_df$label[springer_country_df$year == 2020] = 
    springer_country_df$country[springer_country_df$year == 2020]

background_df = rbind(nature_country_df, 
                           springer_country_df[,colnames(nature_country_df)])
 

ggplot(subset(background_df, country == "United States"), 
       aes(x=as.numeric(year), y=mean,
                          ymin=bottom_CI, ymax=top_CI,
                          fill=corpus)) +
    geom_point() + geom_ribbon(alpha=0.5) + geom_line(alpha=0.5) +
    theme_bw() + 
    xlab("Year of Article") + 
    ylab("Percentage articles with at least one author with US affiliation") +
    ggtitle("Percentage articles with mention/citation of US locations") + 
    scale_fill_brewer(palette="Set2")


ggplot(subset(background_df, country == "United Kingdom"), 
       aes(x=as.numeric(year), y=mean,
                          ymin=bottom_CI, ymax=top_CI,
                          fill=corpus)) +
    geom_point() + geom_ribbon(alpha=0.5) + geom_line(alpha=0.5) +
    theme_bw() + 
    xlab("Year of Article") + 
    ylab("Percentage articles with at least one author with UK affiliation") +
    ggtitle("Percentage articles with mention/citation of UK locations") + 
    scale_fill_brewer(palette="Set2")


ggplot(subset(background_df, country == "People's Republic of China"),
       aes(x=as.numeric(year), y=mean,
                          ymin=bottom_CI, ymax=top_CI,
                          fill=corpus)) +
    geom_point() + geom_ribbon(alpha=0.5) + geom_line(alpha=0.5) +
    theme_bw() + 
    xlab("Year of Article") + 
    ylab("Percentage articles with at least one author with Chinese affiliation") +
    ggtitle("Percentage articles with mention/citation of Chinese locations") + 
    scale_fill_brewer(palette="Set2")


ggplot(subset(background_df, country == "India"), 
       aes(x=as.numeric(year), y=mean,
                          ymin=bottom_CI, ymax=top_CI,
                          fill=corpus)) +
    geom_point() + geom_ribbon(alpha=0.5) + geom_line(alpha=0.5) +
    theme_bw() + 
    xlab("Year of Article") + 
    ylab("Percentage articles with at least one author with Indian affiliation") +
    ggtitle("Percentage articles with mention/citation of Indian locations") + 
    scale_fill_brewer(palette="Set2")


```

Now lets compare nature news citations rate against Springer and Nature articles for a few countries.
We see that the citation rate mostly tracks the Nature article rate.

```{r citation_v_springer_v_nature, out.width="50%", echo=FALSE, warning=FALSE, message=F}


compare_df = rbind(citation_country_df, 
                   springer_country_df,
                   nature_country_df)

ggplot(subset(compare_df, country == "United States"),
       aes(x=as.numeric(year), y=mean,
                          ymin=bottom_CI, ymax=top_CI,
                          fill=corpus)) +
    geom_point() + geom_ribbon(alpha=0.5) + geom_line(alpha=0.5) +
    theme_bw() +
    xlab("Year of Article") +
    ylab("Percentage articles with at least one author with US affiliation") +
    ggtitle("Comparison of US citation rate vs US authorship rate") +
    scale_fill_brewer(palette="Set2")


ggplot(subset(compare_df, country == "United Kingdom"),
       aes(x=as.numeric(year), y=mean,
                          ymin=bottom_CI, ymax=top_CI,
                          fill=corpus)) +
    geom_point() + geom_ribbon(alpha=0.5) + geom_line(alpha=0.5) +
    theme_bw() +
    xlab("Year of Article") +
    ylab("Percentage articles with at least one author with UK affiliation") +
    ggtitle("Comparison of UK citation rate vs UK authorship rate") +
    scale_fill_brewer(palette="Set2")


ggplot(subset(compare_df, country == "People's Republic of China"),
       aes(x=as.numeric(year), y=mean,
                          ymin=bottom_CI, ymax=top_CI,
                          fill=corpus)) +
    geom_point() + geom_ribbon(alpha=0.5) + geom_line(alpha=0.5) +
    theme_bw() +
    xlab("Year of Article") +
    ylab("Percentage articles with at least one author with PRC affiliation") +
    ggtitle("Comparison of PRC citation rate vs PRC authorship rate") +
    scale_fill_brewer(palette="Set2")


ggplot(subset(compare_df, country == "India"),
       aes(x=as.numeric(year), y=mean,
                          ymin=bottom_CI, ymax=top_CI,
                          fill=corpus)) +
    geom_point() + geom_ribbon(alpha=0.5) + geom_line(alpha=0.5) +
    theme_bw() +
    xlab("Year of Article") +
    ylab("Percentage articles with at least one author with Indian affiliation") +
    ggtitle("Comparison of India citation rate vs India authorship rate") +
    scale_fill_brewer(palette="Set2")


```




Now lets compare nature news mentions rate against Springer and Nature articles for a few countries.

```{r four_countries_v_springer_v_nature,  out.width="50%", echo=FALSE, warning=FALSE, message=F}

compare_df = rbind(mention_country_df, 
                   springer_country_df,
                   nature_country_df)

ggplot(subset(compare_df, country == "United States"),
       aes(x=as.numeric(year), y=mean,
                          ymin=bottom_CI, ymax=top_CI,
                          fill=corpus)) +
    geom_point() + geom_ribbon(alpha=0.5) + geom_line(alpha=0.5) +
    theme_bw() +
    xlab("Year of Article") +
    ylab("Percentage articles with at least one mention of the US") +
    ggtitle("Comparison of US mention rate vs US authorship rate") +
    scale_fill_brewer(palette="Set2")


ggplot(subset(compare_df, country == "United Kingdom"),
       aes(x=as.numeric(year), y=mean,
                          ymin=bottom_CI, ymax=top_CI,
                          fill=corpus)) +
    geom_point() + geom_ribbon(alpha=0.5) + geom_line(alpha=0.5) +
    theme_bw() +
    xlab("Year of Article") +
    ylab("Percentage articles with at least one mention of the UK") +
    ggtitle("Comparison of UK mention rate vs UK authorship rate") +

    scale_fill_brewer(palette="Set2")


ggplot(subset(compare_df, country == "People's Republic of China"),
       aes(x=as.numeric(year), y=mean,
                          ymin=bottom_CI, ymax=top_CI,
                          fill=corpus)) +
    geom_point() + geom_ribbon(alpha=0.5) + geom_line(alpha=0.5) +
    theme_bw() +
    xlab("Year of Article") +
    ylab("Percentage articles with at least one mention of the PRC") +
    ggtitle("Comparison of PRC mention rate vs PRC authorship rate") +

    scale_fill_brewer(palette="Set2")


ggplot(subset(compare_df, country == "India"),
       aes(x=as.numeric(year), y=mean,
                          ymin=bottom_CI, ymax=top_CI,
                          fill=corpus)) +
    geom_point() + geom_ribbon(alpha=0.5) + geom_line(alpha=0.5) +
    theme_bw() +
    xlab("Year of Article") +
    ylab("Percentage articles with at least one mention of the India") +
    ggtitle("Comparison of India mention rate vs India authorship rate") +

    scale_fill_brewer(palette="Set2")


```